### 2015-09-16 stuff/dirfields/DirFields.tex
...

@@

%surface (manifold) -> M, S, \Omega 
\newcommand{\M}{M}
%volume element (volume measure) -> \mu, dA, d\Omega
\newcommand{\dA}{dA}
%director field (contra-vector)
\newcommand{\p}{\mathbf{p}}
\newcommand{\q}{\mathbf{q}}
%director field (co-vector, flat of contra-vector)
\newcommand{\pfl}{\mathbf{p}^{\flat}}
\newcommand{\qfl}{\mathbf{q}^{\flat}}
%Frank Oseen Energy (without Lagrange term for normalizing)
\newcommand{\EOS}{E_{\text{OS}}}
%Normalizing energy
\newcommand{\EN}{E_{n}}
%Laplace-Beltrami or Rot-Rot-Laplace
\newcommand{\LB}{\Delta_{\text{\tiny RR}}}
%Laplace-CoBeltrami or Grad-Div-Laplace
\newcommand{\LCB}{\Delta_{\text{\tiny GD}}}
%Laplace-deRham
\newcommand{\LDR}{\Delta_{\text{\tiny dR}}}
%Landau symbol
\renewcommand{\O}{\mathcal{O}}

@@

>>>>

\usepackage{hyperref}

@@

%surface (manifold) -> M, S, \Omega 
\newcommand{\M}{M}
%volume element (volume measure) -> \mu, dA, d\Omega
\newcommand{\dA}{dA}
%director field (contra-vector)
\newcommand{\p}{\mathbf{p}}
\newcommand{\q}{\mathbf{q}}
%director field (co-vector, flat of contra-vector)
\newcommand{\pfl}{\mathbf{p}^{\flat}}
\newcommand{\qfl}{\mathbf{q}^{\flat}}
%discrete director field (co-vector, flat of contra-vector)
\newcommand{\pflh}{\mathbf{p}^{\flat}_{h}}
%discrete director field (contra-vector)
\newcommand{\ph}{\mathbf{p}_{h}}
%discrete PD director field (co-vector, flat of contra-vector)
\newcommand{\PDpflh}{\underline{\mathbf{p}}^{\flat}_{h}}
%discrete director field (contra-vector)
\newcommand{\PDph}{\underline{\mathbf{p}}_{h}}
%Frank Oseen Energy (without Lagrange term for normalizing)
\newcommand{\EOS}{E_{\text{OS}}}
%Normalizing energy
\newcommand{\EN}{E_{n}}
%Laplace-Beltrami or Rot-Rot-Laplace
\newcommand{\LB}{\Delta^{\text{\tiny RR}}}
%Laplace-CoBeltrami or Grad-Div-Laplace
\newcommand{\LCB}{\Delta^{\text{\tiny GD}}}
%Laplace-deRham
\newcommand{\LDR}{\Delta^{\text{\tiny dR}}}
%discrete Laplace-Beltrami or Rot-Rot-Laplace
\newcommand{\LBh}{\LB_{h}}
%Laplace-CoBeltrami or Grad-Div-Laplace
\newcommand{\LCBh}{\LCB_{h}}
%Laplace-deRham
\newcommand{\LDRh}{\LDR_{h}}
%Landau symbol
\renewcommand{\O}{\mathcal{O}}
%Vertices
\newcommand{\V}{\mathcal{V}}
%Edges
\newcommand{\E}{\mathcal{E}}
%Faces
\newcommand{\F}{\mathcal{F}}
%edge vector
\newcommand{\e}{\mathbf{e}}
%vertex vector
\renewcommand{\v}{\mathbf{v}}

@@

\section{A DEC approach}

For further information see for example \cite{whitney,hirani}.

  \subsection{Surface Mesh}
    ...wellcentered manifoldlike simplicial complex, bla, bla, blub...

  \subsection{Discrete 1-forms}

  The main concept to represent a discrete 1-form \( \pflh\in\Lambda_{h}(K) \) is to approximate the contraction of the continuous 1-Form 
  \( \pfl\in\Lambda(M) \) on all edges \( e\in\E \)
  \begin{align}
    \pflh(e) := \int_{\pi(e)}\pfl \approx \int_{0}^{1} \pfl_{X_{e}(\tau)}\left( \dot{X}_{e}(t) \right) dt = \pfl_{X_{e}(\tau)}(\e)
  \end{align}
  where \( \pi:K\rightarrow \M \) is the glueing map, who project the elements of the surface mesh to the manifold.
  \( X_{e}(t) = t\v_{2} + (1-t)\v_{1} \) is the linear barycentric parametrisation of the edge \( e=\left[ v_{1}, v_{2} \right]\).
  The existence of a intermediate value \( \tau\in\left[ 0,1 \right] \), so that \( \e\in T_{X_{e}(\tau)}M \),
  is ensured by the mean value theorem.
  Other discrete forms of arbitrary degree and theirs hodge duals can be interpreted in a similarly way.

  \subsection{Discrete Laplace operators}

    In the discrete exterior calculus discrete Operators are defined by successively interpretation of the basic operations on the forms, 
    like the Hodge operator \( * \) or the exterior derivative \( \exd \), as geometric operators on the simplices, 
    like the Voronoi dual operator \( \star \) or the boundary operator \( \partial \) (see \cite{hirani}).
    This results for example to a discrete definition of \( \LB \) for a discrete 1-form \( \pflh\in\Lambda^{1}_{h}(M) \) on a edge 
    \(e\in\E \)
    \begin{align}
      \LBh\pflh(e) &:=  \left(*\exd*\exd \pflh\right)(e)
                     = -\frac{\left| e \right|}{\left| \star e \right|} \left(\exd * \exd  \pflh\right)(\star e) \\
                    &=  -\frac{\left| e \right|}{\left| \star e \right|} \left(* \exd  \pflh\right)(\partial\star e)
                     = -\frac{\left| e \right|}{\left| \star e \right|} \sum_{f\succ e} s_{f,e} \left(* \exd \pflh\right) (\star f) \\
                    &= -\frac{\left| e \right|}{\left| \star e \right|} \sum_{f\succ e} \frac{s_{f,e}}{\left| f \right|} 
                                \left(\exd\pflh\right)(f)
                     = -\frac{\left| e \right|}{\left| \star e \right|} \sum_{f\succ e} \frac{s_{f,e}}{\left| f \right|} \pflh(\partial f)\\
                    &= -\frac{\left| e \right|}{\left| \star e \right|} \sum_{f\succ e} \frac{s_{f,e}}{\left| f \right|}
                                \sum_{\tilde{e}\prec f} s_{f,\tilde{e}} \pflh(\tilde{e})
    \end{align}
    or for \( \LCB \)
    \begin{align}
      \LCBh\pflh(e) &:= \left( \exd * \exd * \pflh\right) (e)
                      = \left( * \exd *\pflh \right) (\partial e) \\
                     &= \sum_{v\prec e} s_{v,e} \left( * \exd * \pflh \right)(v) 
                      = \sum_{v\prec e} \frac{s_{v,e}}{\left| \star v \right|} \left( \exd * \pflh \right)(\star v)\\
                     &= \sum_{v\prec e} \frac{s_{v,e}}{\left| \star v \right|} \left( * \pflh \right)(\partial\star v)
                      = -\sum_{v\prec e} \frac{s_{v,e}}{\left| \star v \right|} \sum_{\tilde{e}\succ v}
                                s_{v,\tilde{e}} \left( * \pflh \right) (\star \tilde{e}) \\
                     &= -\sum_{v\prec e} \frac{s_{v,e}}{\left| \star v \right|} \sum_{\tilde{e}\succ v}
                                s_{v,\tilde{e}} \frac{\left| \star\tilde{e} \right|}{\left| \tilde{e} \right|}\pflh(\tilde{e})
    \end{align}
    where the sign \( s_{f,e} \) is +1 if the face \( f\succ e \) is in the left of the edge \( e \) (-1 otherwise)
    and \( s_{v,e} \) is +1 if the edge \( e\succ v \) points to the vertex \( v \) and -1 if \( e \) points away from \( v \).
    We see \( \LBh \), \( \LCBh \) and also \( \LDRh := -\LBh -\LCBh \) are linear operators in \( \pflh(\tilde{e}) \)
    and therefor results in sparse matrices if the \( \pflh(\tilde{e}) \) are our degree of freedoms.
    ...explainable figures...

  \subsection{Discrete norm}

    Approximating the norm \( \left\| \pfl \right\| \) on a edge \( e\in\E \) is not so easy like the development of discrete linear
    operators. 
    We only know how \( \pflh \) "act" on a single edge, so \( \pflh \) gives us only one dimensional informations.
    In other words, we only know the proportion \( \ph \cdot \e = \pflh (e)\) of the discrete contra vector field \( \ph = \left( \pflh \right)^{\sharp} \) in the
    \( \e \) direction, but we don't know the length of \( \ph \) defined on this edge.

    ...some averaging techniques and why these sucks...

    One way out is to rise the dimension of the discrete 1-forms, therefor we introduce some bases at the intersection 
    \( c(e) = e\cap (\star e) \) of a edge and its dual edge to describe discrete contra- and covariant vector fields in a local (flat)
    coordinate system.

    The basis for contravariant vectors at \( c(e) \) is composed of
    \begin{align}
      \partial_{e}X &:= \e = \v_{2} - \v_{1} \\
      \partial_{\star e}X &:= *\e =  c(f_{2}) - c(f_{1})
    \end{align}
    if \( e=\left[ v_{1}, v_{2} \right] \) and the face \( f_{1} \) lay right and \( f_{2} \) left of \( e \) (...figure...). 
    This definitions are consistent with the canonical basis, if the position \( X \) is a barycentric parametrisation of the edge \( e \)
    resp. its dual, i.e. for
    \begin{align}
      X_{e}(t_{e}) &= t_{e}\v_{2} + \left( 1 - t_{e} \right)\v_{2} \text{ for } t\in[0,1]\\
      X_{\star e}(t_{\star e}) &=\begin{cases}
                        2t_{\star e}c(e) + \left( 1 - 2t_{\star e} \right) c(f_{1}) & \text{if } t_{\star e}\in\left[ 0 , \frac{1}{2} \right] \\
                        \left(1 - t_{\star e}\right) c(e) + \left( 2t_{\star e} - 1 \right) c(f_{1}) 
                                & \text{if } t_{\star e}\in \left[\frac{1}{2} , 1\right]
                      \end{cases}
    \end{align}
    holds \( \partial_{e}X = \partial_{t_{e}}X_{e} \) 
    and \( \partial_{\star e}X = \frac{1}{2} \left( \partial_{t_{\star e}}X_{\star e}|_{\left[ 0 , \frac{1}{2} \right]} 
                                                   + \partial_{t_{\star e}}X_{\star e}|_{\left[\frac{1}{2}, 1 \right]}\right) \).
    Therefore we get the local metric tensor
    \begin{align}
      g_h(e) &= \left| e \right|^{2} \left( dx^{e} \right)^{2} + \left| \star e \right|^{2} \left( dx^{\star e} \right)^{2}
    \end{align}
    where \( \left\{  dx^{e}, dx^{\star e} \right\} \) are the dual base of \( \left\{ \partial_{e}X, \partial_{\star e}X \right\} \),
    i.e. \( dx^{i}\left(\partial_{j}X\right) = \delta^{i}_{j} \) for \( i,j\in\left\{ e, \star e \right\} \).
    This gives us the great possibility to define
    \begin{align}
      \PDpflh(e) &:= \pflh(e)dx^{e} + \pflh(\star e)  dx^{\star e} \\
                   &= \pflh(e)dx^{e} - \frac{\left| \star e \right|}{\left| e \right|} \left(*\pflh\right)(e)  dx^{\star e} \\
      \PDph(e) &:= \left(\PDpflh(e)\right)^{\sharp} 
                 = \frac{1}{\left| e \right|^{2}} \pflh(e)\partial_{e}X + \frac{1}{\left|\star e \right|^{2}} \pflh(\star e)\partial_{\star e}X \\
               & =  \frac{1}{\left| e \right|} \left( \frac{1}{\left| e \right|} \pflh(e)\partial_{e}X 
                                                     -\frac{1}{\left|\star e \right|} \left(*\pflh\right)(e) \partial_{\star e}X\right)
    \end{align}
    We call \( \PDpflh(e) \) the discrete Primal-Dual-1-form (short PD-1-form) and \( \PDph \) the discrete Primal-Dual-vector field
    (short PD-vector field).
    Hence, we get for the square of the norm on \( c(e) \) by contract the PD-1-Form with its corresponding PD-Vector
    \begin{align}
      \left\| \PDpflh \right\|^{2}(e) := \left( \PDpflh\left( \PDph \right) \right)(e)
                                       = \frac{1}{\left| e \right|^{2}}\left( \left[ \pflh(e) \right]^{2} 
                                                                             +\left[ \left(*\pflh\right)(e) \right]^{2}\right)
    \end{align}

\bibliography{bibl}
\bibliographystyle{alpha}
